Verifying Spatial Properties of Array Computations

### 2017-04-17 oopsla17/algorithms.tex
Second part of inference

>>>>

Now we convert coalesced indexing schemes into specifications. This happens in
three stages. First, we see if there is an all infinite indexing scheme. Second,
we determine if the indexing schemes are interval schemes so that they can be
represented exactly and if not how it can be altered into interval schemes to be
represented as approximations. Third, we split up the interval schemes so that
the region constants can express them. Finally, we convert the interval schemes
into a joins of meets of region constants.

If there is all infinite indexing scheme, then we cannot represent it
syntactically. This happens when use of constant indices in one or more array
terms causes the offsets to be subsumed. The simplest case is when there is an
array term with all of its indices constant.

In our example, there are no constant indices, so we proceed.

After handling the trivial case, correct approximation is chosen. Recall
\Cref{def:interval-scheme}, an indexing scheme is an interval scheme if each of
the sets in products are holed intervals. If that is the case, we trivially
convert the sets in each dimension of indexing schemes to holed interval
representation and proceed to the next stage and output an exact specification
at the end. If not, then the output will definitely have an upper bound and
potentially a lower bound.

Upper bound is established by \emph{elongating} index schemes in multiple
dimensions such that they become interval schemes. The elongation function is
given as
%
\begin{align*}
  \mathit{elongate}([a,b]) = \begin{cases}
    \interv{0}{b}{\textsf{false}} & a > 1 \\
    \interv{a}{0}{\textsf{false}} & b < -1
  \end{cases}
\end{align*}

Since index schemes of the previous stage are all contiguous and form closed
intervals, 0 lying between the lower and upper bounds of intervals is sufficient
to form an interval scheme. If any dimension of an index scheme is elongated,
then the whole index scheme is elongated. In the initial set of index schemes,
all schemes that were already interval schemes form a lower bound, while those
schemes and the new interval schemes generated by elongation form the upper
bound. The reason a lower bound may not be generated is all indexing schemes
with might need elongation.

The index schemes in the example are all interval schemes, so we proceed to the
next stage noting the output will be exact.

There is no direct translation from an interval scheme such as $\interv{-2}{1}{}
\times \interv{-1}{1}{}$ to region constants. The reason is region constants
\texttt{forward} and \texttt{backward} require upper and lower bounds to be 0
respectively and \texttt{centered} requires the upper bound to be the negation
of the lower bound. The solution is to use \Cref{lem:vector-union} to break up
interval schemes into two producing $\interv{-2}{0}{} \times \interv{-1}{1}{}$
and $\interv{0}{1}{} \times \interv{-1}{1}{}$.

The extended example does not have such representation problem, so the interval
schemes are left as they are.

The final stage of inference is simply mapping interval schemes into region
constants. Each interval scheme is connected with \texttt{+} and each dimension
of the interval scheme is connected with \texttt{*}. The intervals at each
dimension are mapped to region constants using the bijection defined in
\Cref{sec:semantics}.

In the example, $[-1, 1] \times [0, 0]$ is mapped to \texttt{\stencil{c}{1}{1}{}
* \stencil{p}{2}{}{}} and $[0, 0] \times [-1, 1]$ is mapped to
\texttt{\stencil{p}{1}{}{} * \stencil{c}{2}{1}{}}, producing the overall
specification:
%
\begin{equation*}
  \texttt{stencil readOnce,} \;
  \stencil[s]{c}{1}{1}{} \texttt{*} \stencil[s]{p}{2}{}{} \; \texttt{+} \;
  \stencil[s]{p}{1}{}{}  \texttt{*} \stencil[s]{c}{2}{1}{}
\end{equation*}
